Hydrodynamics of a Semipermeable Vesicle Under Flow and Confinement

### 2020-12-29 vesicle_code/docs/permeableVes/main.tex
added a todo
editing early sections

%\title{Semi-Permeable Vesicles: Mathematical Formulation, Numerics, and
%Physics}
\title{Hydrodynamics of a  Semi-permeable Vesicle Under Flow and Confinement}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a]{Bryan Quaife}
\author[a]{Ashley Gannon}
\author[b]{Howard A. Stone}
\author[c,1]{Y.-N. Young}

@@

\begin{abstract}
In this work we use boundary integral simulations to investigate the effects of water permeability on the hydrodynamics of a vesicle under various flows and confinement configurations.
Using the Helfrich free energy extended for the coupling of membrane permeability with the membrane elastic stress, tension and mechanical stress, we illustrate that, in the absence of an osmotic stress gradient, a typical liposome may be significantly affected by membrane permeability over a sufficiently long time or under strong confinement. 
Lipid bilayer membranes are permeable to water molecules, and water influx/efflux  is often assumed to be negligible in the absence of osmolarity, even when the membrane undergoes large deformation. Our simulation results quantify the conditions for such assumption to be valid, and shed light on how confinement can be utilized to estimate membrane permeability.
%
%osmotic stress gradient 
%We describe a mathematical model for a semi-permeable vesicle and derive
%  a boundary integral equation formulation for simulating its
%  hydrodynamics.  
%
%Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.
\end{abstract}

@@

\dropcap{W}ater exchange is essential for a living cell to adapt to its environment over a wide range of timescales \cite{CadartVenkovaRechoEtAl2019_NaturePhys,AlbertsMolecularBiology,YangMaVerkman2001_JBC,SugieIntaglietta2018_AmJPhysiolHCP,SaadounPapadopoulosWatanabeEtAl2005_JCS,Verkman2008_JMM,BerthaudEtAl2016_SM,Keren2011_EurBJ,TaloniKardashSalmanEtAl2015_PRL}.  
%
%Water permeation across the cell membrane due to an osmotic shock has been well studied \cite{AlbertsMolecularBiology}. 
Recent findings confim that both osmotic and mechanical stresses contribute to significant water permeation for cells to migrate under strong confinement \cite{JiangSun2013_BJ,StrokaJiangChenEtAl2014_Cell,LiMoriSun2015_PRL,yao-mor2017}, leading to a dynamic surface-to-volume ratio of a migrating cell. 
While stress-induced release of messengers in cells has been well studied \cite{Wan2008_PNAS, ForsythWan2011_PNAS,Russell-PuleriPazAdams2016_AJHCP,ZhangShenHoganBarakatMisbah2018_BJ,GordonShimmelFrye2020_FP},
%osmotic effects on cellular dynamics have been well studied \cite{AlbertsMolecularBiology}, 
mechanically induced water permeation is often associated with membrane poration or rupture under extreme stresses \cite{HarmanBertrandJoos2017_CJP,RazizadehNikfarPaulLiu2020_BJ}. In this work we use modeling and direct numerical simulations to show that, 
over the appropriate time scales or under strong confinement,
the intrinsic membrane permeability to water can give rise to significantly distinct hydrodynamics of a semipermeable vesicle without poration or rupture in the lipid bilayer membrane. 
%
%
%One of the most direct ways to induce water permeation across cell membranes is to impose an osmotic shock }.
%
%A gradient in the osmotic stress across the cellular membrane is the most direct way to induce water permeation, yet 
%%
%%In response to external stimuli, cells move, deform, and release signaling molecules, such as ATP in red blood cells \cite{ZhangShenHoganBarakatMisbah2018_BJ} and cyclooxygenase-2 in endothelial cells \cite{Russell-PuleriPazAdams2016_AJHCP}. 
%%
%%Water permeation is often neglected in such release processes, yet 
%%
%recent findings suggest that both osmotic and mechanical stresses can induce water permeation that is significant enough for tumor cells to migrate under strong confinement \cite{StrokaJiangChenEtAl2014_Cell,LiMoriSun2015_PRL,yao-mor2017}, leading to a dynamic surface-to-volume ratio over time. 
%
%
%To elucidate the effects of mechanically induced water permeation (which is a great challenge to quantify in experiments), in this work we use accurate numerical simulations to study the hydrodynamics of a semi-permeable vesicle in various configurations.

@@

The lipid bilayer membrane is permeable to water and water-soluble molecules \cite{Dick1964_JTB,FettiplaceHaydon1980_PhysRev,DeamerBramhall1986_ChemPhysLipids,Grafmueller2019_ABLS}. 
%
%In the presence of an osmotic gradient across the membrane (such as a contrast in the concentration of an impermeable solute), water flows against the osmotic gradient  to neutralize the osmolarity. 
%
Without aquaporins or other membrane channels that promotes water permeability, the intrinsic water permeability is known to depend on the temperature \cite{OlbrichRawiczNeedhamEtAl2000_BJ}, lipid composition \cite{OlbrichRawiczNeedhamEtAl2000_BJ}, and the hydration of the membrane \cite{MarrinkBerendsen1994_JPhysChem}. In the red blood cell membrane, aquaporin-1 channels give rise to a much higher water permeability of $\sim 1.8\times 10^{-2}$ cm/s in erythrocytes \cite{YangMaVerkman2001_JBC} while the intrinsic water permeability of a lipid bilayer membrane is in the range of $10^{-4}-10^{-3}$ cm/s \cite{ThompsonHuang1966_ANYAS,FettiplaceHaydon1980_PhysRev,Grafmueller2019_ABLS,Dimova2020_GVB,BhatiaRobinsonDimova2020_SoftMatt}. 
%
%The swelling/shrinking of a cell is inevitably coupled to changes in membrane tension and deformation. Under a strong confinement (such as going through a microcapillary or submicron slit),  the transmembrane water flow may have comparable contributions from both osmotic and mechanical stresses \cite{StrokaJiangChenEtAl2014_Cell,LiMoriSun2015_PRL,yao-mor2017}.  
Molecular dynamic (MD) 
%and coarse-grained (CG) MD 
simulations show that a stretched RBC membrane porates when the membrane tension reaches the order of $2-4$ mN/m (\cite{RazizadehNikfarPaulLiu2020_BJ} and references therein), giving rise to enhanced membrane permeability to both water and macromolecules. Consequently, as long as the lipid bilayer membrane remains intact (no poration or rupture), RBCs and vesicles are assumed to be of the same volume throughout experiments of durations no longer than tens of minutes. 
%
%membrane poration is inevitable when membrane tension is of the order of $2-4$ mN/m (\cite{RazizadehNikfarPaulLiu2020_BJ} and references therein). 
%%The CGMD results further show that a stretched RBC membrane may have pores that might contribute to transmembrane transport of both water and ATP molecules \cite{RazizadehNikfarPaulLiu2020_BJ}. 
%Thus an important outstanding question in stress-induced ATP release of highly deformed RBCs is if there is any transmembrane flow when RBCs release ATP as they experience large
%mechanical load due to confinement or large shear stress?  To address such question, the first step we take is to focus on the permeable water flux induced by mechanical stress.
%%
%%RBCs are known to release ATP under large deformation and microfluidic experiments have been used to study the correlation between RBC deformation and ATP release \cite{Wan2008_PNAS,ForsythWan2011_PNAS,ZhangShenHoganBarakatMisbah2018_BJ}.
%%
%%Why is it interesting to study the dynamics of a semi-permeable membrane induced by mechanical stress/confinement?
%%
%%The osmolarity-induced water flux has been well studied, while mechanical pressure may also contribute to the permeable water flow \cite{yao-mor2017}.
%%The response of a cell to an osmotic gradient is inevitably both mechanical (tension redistribution and shape change) and chemical (water flux across the membrane).  
Under what condition is this a good assumption? Do vesicles conserve volume over hours under flow? What is the dynamic consequence when water influx/efflux affects the membrane hydrodynamics?
To answer these questions, we investigate the balance between hydrodynamics, tension and elasticity of a semipermeable membrane under various flowing conditions and confinement. 
%Specifically our numerical simulations yield the quantification of the time scale and length scale over which water permeation becomes important.

% In this work we use numerical simulations to examine the hydrodynamics of a semipermeable vesicle in two dimensions. 
In microfluidic experiments, it is challenging to keep a freely suspended RBC or vesicle in a steady flow for more than a few hours. It is also difficult to quantify the vesicle area/volume ratio accurately. Using numerical simulations we can quantify  the effects of water permeability on the hydrodynamics of a vesicle under mechanical stresses. 
Vesicles (self-enclosed lipid bilayer membranes) have been used as a model system to study reshaping, remodeling and scission of cell membranes due to osmotic stress \cite{OgleckaEtAl2014_eLife,CamposSaric2020_bioRxiv,Dimova2020_GVB,BhatiaChrist2020_SoftMatt}. Under an osmotic shock, a vesicle may undergo oscillatory variation in its enclosed volume \cite{ChabanonHoLiedberg2017_BJ}. When trapped in a microfluidic channel, an osmotic shock induces 
 vesicle shape changes, tank-treading and tumbling motions \cite{BhatiaRobinsonDimova2020_SoftMatt,BhatiaChrist2020_SoftMatt}, an illustration that the permeable water flow is inevitably coupled with the mechanical pressure jump across the membrane\cite{yao-mor2017}, which could become comparable to osmotic stress when the membrane deformation is large \cite{LiMoriSun2015_PRL,yao-mor2017}.  In the absence of an osmotic gradient, we first show that a freely suspended semipermeable vesicle behaves differently from an impermeable vesicle over a long time. Next we show that, at normal physiological conditions, the permeable water flow can be amplified by extreme confinement and results in observable change in both vesicle hydrodynamics and reduced area.  These results will shed light on the dynamic aging process of living cells that constantly go through flows and constrictions over a period of days.

%
%\paragraph{Literature to maybe cite}
%\cite{ali-ovr-ran2020} show how beading can occur on nanotubes. This
%happens in regions where proteins congregate and this can lead to
%semi-permeability?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]
  \centering
  \includegraphics[width=0.5\textwidth]{figures/schematic.pdf}
  \caption{\label{fig:sketch} Schematics showing the effects of
  semipermeability on vesicle hydrodynamics in various configurations.
  The slipper in (a) is the steady-state shape of a semi-permeable
  vesicle in a Poiseuille flow.}
\end{figure}

\section*{Formulation}
We consider a two-dimensional semi-permeable vesicle (permeable only to water) suspended in a viscous fluid, see Figure~\ref{fig:sketch}.
To focus on the effect of water flow on vesicle hydrodynamics in various conditions, we assume the fluid is the same inside and outside the vesicle membrane.
%($\gamma$) with an outward normal $\nn$.
On the membrane the permeable water flux gives rise to a difference between the fluid velocity ${\bf u}({\bf x}\in \gamma)$
and the membrane velocity $\dot{\bf x}$ \cite{yao-mor2017}:
\begin{align}
  \label{eqn:vesicleFlux}
  \uu - \dot{\xx} = - k_w (R T \triangle c + \ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $\gamma$ denotes the membrane boundary with an outward normal $\nn$ (Figure~\ref{fig:sketch}), $k_w$ is the water permeability, $R$ is the ideal gas constant, $T$ is the temperature ($^{\circ}$K), $\triangle c$ is contrast in a solute concentration (mol/volume),
and $\ff_\mathrm{mem}$ is the membrane stress that consists of both bending $ \ff_\mathrm{ben}= -k_b \xx_{ssss}$ and tension $\ff_\mathrm{ten}=(\sigma \xx_s)_s$ ($s$
is the arclength along the membrane, see Supplementary Material). Using $k_w\sim 10^{-12}$ m$^2$s/kg (typical of mammalian cell membranes \cite{LiMoriSun2015_PRL}) and an osmolarity $\triangle c = 0.1$ mol/L (consistent with the osmotic forces in vesicle aspiration experiments \cite{OlbrichRawiczNeedhamEtAl2000_BJ}), the corresponding water flow is of the order $v_\mathrm{o}=0.1 \mu$m/s. Assuming that the membrane bending stiffness $k_b = 10^{-19}$J, a membrane tension $10^{-3}$N/m (one tenth of the lysis tension) and a characteristic length scale of $10^{-6}$m, the tension contribution ($v_\mathrm{t}$) and elastic stress contribution 
($v_\mathrm{e}$) to the permeable water flow can be scaled to $v_\mathrm{o}$ as $v_\mathrm{o}:v_\mathrm{t}:v_\mathrm{e}=1:10^{-2}:10^{-6}$.  Furthermore, we note that in the absence of an osmotic stress, the water permeable flow can be significant when the membrane tension is large due to the external stress such as the hydrodynamics stress due to an external flow and/or due to confinement.
In the following we will explore both.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Effects of membrane water permeability}
We consider a single semi-permeable vesicle in familiar configurations to highlight the effects of membrane water permeability.
In free space we put a semi-permeable vesicle in a quiescent flow, a planar shear flow and a Poiseuille flow.
We also consider a semi-permeable vesicle in confinement: a long channel and a contracting channel.
For each example, we fix the vesicle length and examine the enclosed area as a dynamical consequence of water flow, vesicle shape deformation and membrane tension distribution.

In experiments vesicle under shear flows are measured for a minute (chapter 19 in \cite{Dimova2020_GVB}), a duration over which vesicle can be assumed impermeable \cite{AbkarianViallat2005_BJ}. Our simulations of a semi-permeable vesicle
in these flows show that water permeability may lead to different vesicle hydrodynamics over a time scale of 10 to 100 minutes. 
%{\bf YNY: is this correct? The shear flow results seem to suggest that the permeability may lead to change in area in less than a second???}
For vesicles in a narrow micro-fluidic channel, the typical time scale is of $ms$ \cite{abk-fai-sto2006}. Our simulations show that strong confinement can lead to amplification of water permeability at such short time scale.
%
%all vesicles have a length
%$L$, but the area is not conserved because of semi-permeability. We vary
%the initial vesicle reduced area, the permeability rate, and the flow
%rate. For all simulations, the bending modulus is $10^{-19}J$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/relaxationComposite.tikz}
  \else
  \includegraphics{figures/relaxationComposite.pdf}
  \fi
  \caption{\label{fig:relaxationComposite} (a): Snapshots of four
  semi-permeable vesicles with $\beta = 10^{-3}$ submerged in a
  quiescent flow. The fluid flux along the red regions is into the
  vesicle, and the fluid flux along the blue regions is out of the
  vesicle. In all cases, the vesicle reaches a steady state circular
  shape. (b): The reduced area of each of the simulations. The
  particular snapshots in (a) occur at the marks along the curve.}
\end{figure*}
In a quiescent flow, a semipermeable vesicle in free space can be quantified in terms of its area dynamics and its transition to the equilibrium shape:
Our analysis in Supplementary Material shows that a semi-permeable vesicle
relaxes to a circle at a rate that is proportional to $\beta$, and the
steady state tension is $\Lambda = 2\pi^2/L^2$. Our analysis further shows that the time for a semipermeable
vesicle to deviate from its initial reduced area and to reach its steady
state is proportional to $\beta^{-1}$. For this reason, our numerical
results below are reported in terms of the scaled time $\beta t$.

%This is verified for a
%vesicle with initial reduced area $\nu = 0.65$ and with different
%semi-permeability rates. In Table~\ref{tbl:ellipseRelaxTimes} we report
%the required time, $t_\mathrm{steady}$, for the vesicle's reduced area
%to be within 1.3\% of its steady state reduced area, and the required
%time, $t_\mathrm{inflate}$, for the vesicle's reduced area to exceed
%1.3\% of its initial reduced area. Both times are proportional to
%$\beta^{-1}$, and for this reason, our numerical results are reported in
%terms of the scaled time $\beta t$.
%\begin{table}[htp]
%  \centering
%  \caption{\label{tbl:ellipseRelaxTimes} The time for a semi-permeable
%  vesicle to its steady state ($t_\mathrm{steady}$) and to begin
%  inflating ($t_\mathrm{inflate}$).}
%  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
%    $\beta$ & $t_\mathrm{steady}$  & ratio 
%            & $t_\mathrm{inflate}$ & ratio \\
%    \midrule
%    $10^{0}$  & $6.45 \times 10^{0}$ & ---  
%              & $2.17 \times 10^{-2}$ & --- \\ 
%    $10^{-1}$ & $2.00 \times 10^{1}$ & 3.10 & 
%                $1.25 \times 10^{-1}$ & 5.76 \\
%    $10^{-2}$ & $9.61 \times 10^{1}$ & 4.81 & 
%                $1.59 \times 10^{0}$ & 12.72 \\
%    $10^{-3}$ & $8.36 \times 10^{2}$ & 8.71 & 
%                $2.19 \times 10^{1}$ & 13.77 \\
%    $10^{-4}$ & $8.15 \times 10^{3}$ & 9.74 & 
%                $2.14 \times 10^{2}$ & 9.77 \\
%    $10^{-5}$ & $8.03 \times 10^{4}$ & 9.85 & 
%                $2.20 \times 10^{3}$ & 10.14 \\
%    \bottomrule
%  \end{tabular}
%\end{table}

@@

Between $t=10$ seconds and $t=10$ minutes, the 
intermediate shapes of semi-permeable
vesicles resemble the shape of an impermeable vesicle. This can be
explained by considering the two terms that contribute to the vesicle
velocity: (i) the permeability velocity, $\beta(\ff \cdot \nn) \nn$, and
(ii) the force balance velocity, $\SS[\ff]$. Since $\beta = 10^{-6}$,
the permeability velocity is much smaller than the force balance
velocity at early times, and the semi-permeable vesicle dynamics
resembles those of an impermeable vesicle. As the force balance velocity
decreases, the permeability velocity dominates, and the vesicle begins
to inflate to a circle with no vanishing force and velocity. 
%{\bf YNY:We should put a tick on the time when this transition occurs for each of the vesicles}

>>>>

%\title{Semipermeable Vesicles: Mathematical Formulation, Numerics, and
%Physics}
\title{Hydrodynamics of a  Semipermeable Vesicle Under Flow and
Confinement}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a]{Bryan Quaife}
\author[a]{Ashley Gannon}
\author[b]{Howard A.~Stone}
\author[c,1]{Y.-N.~Young}

@@

\begin{abstract}
In this work we use boundary integral simulations to investigate the
effects of water permeability on the hydrodynamics of a vesicle under
various flows and confinement configurations. Using the Helfrich free
energy extended for the coupling of membrane permeability with the
membrane elastic stress, tension and mechanical stress, we illustrate
that, in the absence of an osmotic stress gradient, a typical liposome
may be significantly affected by membrane permeability over a
sufficiently long time or under strong confinement. Lipid bilayer
membranes are permeable to water molecules, and water influx/efflux is
often assumed to be negligible in the absence of osmolarity, even when
the membrane undergoes large deformation. Our simulation results
quantify the conditions for such assumption to be valid, and shed
light on how confinement can be utilized to estimate membrane
permeability.
%
%osmotic stress gradient 
%We describe a mathematical model for a semipermeable vesicle and derive
%  a boundary integral equation formulation for simulating its
%  hydrodynamics.  
%
%Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.
\end{abstract}

@@

\dropcap{W}ater exchange is essential for a living cell to adapt to its
environment over a wide range of
timescales~\cite{CadartVenkovaRechoEtAl2019_NaturePhys,
AlbertsMolecularBiology, YangMaVerkman2001_JBC,
SugieIntaglietta2018_AmJPhysiolHCP,
SaadounPapadopoulosWatanabeEtAl2005_JCS, Verkman2008_JMM,
BerthaudEtAl2016_SM, Keren2011_EurBJ, TaloniKardashSalmanEtAl2015_PRL}.  
%
%Water permeation across the cell membrane due to an osmotic shock has been well studied \cite{AlbertsMolecularBiology}. 
Recent findings confirm that both osmotic and mechanical stresses
contribute to significant water permeation for cells to migrate under
strong confinement~\cite{JiangSun2013_BJ, StrokaJiangChenEtAl2014_Cell,
LiMoriSun2015_PRL, yao-mor2017}, leading to a dynamic surface-to-volume
ratio of a migrating cell. While stress-induced release of messengers in
cells has been well studied~\cite{Wan2008_PNAS, ForsythWan2011_PNAS, 
Russell-PuleriPazAdams2016_AJHCP, ZhangShenHoganBarakatMisbah2018_BJ,
GordonShimmelFrye2020_FP},
%osmotic effects on cellular dynamics have been well studied
%\cite{AlbertsMolecularBiology}, 
mechanically induced water permeation is often associated with membrane
poration or rupture under extreme
stresses~\cite{HarmanBertrandJoos2017_CJP,
RazizadehNikfarPaulLiu2020_BJ}. In this work we use modeling and direct
numerical simulations to show that, over the appropriate time scales or
under strong confinement, the intrinsic membrane permeability to water
can give rise to significantly distinct hydrodynamics of a semipermeable
vesicle without poration or rupture in the lipid bilayer membrane. 
%
%
%One of the most direct ways to induce water permeation across cell membranes is to impose an osmotic shock }.
%
%A gradient in the osmotic stress across the cellular membrane is the most direct way to induce water permeation, yet 
%%
%%In response to external stimuli, cells move, deform, and release signaling molecules, such as ATP in red blood cells \cite{ZhangShenHoganBarakatMisbah2018_BJ} and cyclooxygenase-2 in endothelial cells \cite{Russell-PuleriPazAdams2016_AJHCP}. 
%%
%%Water permeation is often neglected in such release processes, yet 
%%
%recent findings suggest that both osmotic and mechanical stresses can induce water permeation that is significant enough for tumor cells to migrate under strong confinement \cite{StrokaJiangChenEtAl2014_Cell,LiMoriSun2015_PRL,yao-mor2017}, leading to a dynamic surface-to-volume ratio over time. 
%
%
%To elucidate the effects of mechanically induced water permeation (which is a great challenge to quantify in experiments), in this work we use accurate numerical simulations to study the hydrodynamics of a semipermeable vesicle in various configurations.

@@

The lipid bilayer membrane is permeable to water and water-soluble
molecules~\cite{Dick1964_JTB, FettiplaceHaydon1980_PhysRev,
DeamerBramhall1986_ChemPhysLipids, Grafmueller2019_ABLS}. 
%
%In the presence of an osmotic gradient across the membrane (such as a contrast in the concentration of an impermeable solute), water flows against the osmotic gradient  to neutralize the osmolarity. 
%
Without aquaporins or other membrane channels that promote water
permeability, the intrinsic water permeability is known to depend on the
temperature~\cite{OlbrichRawiczNeedhamEtAl2000_BJ}, lipid
composition~\cite{OlbrichRawiczNeedhamEtAl2000_BJ}, \todo[inline]{Same
reference twice?} and the hydration of
the membrane~\cite{MarrinkBerendsen1994_JPhysChem}. In the red blood
cell membrane, aquaporin-1 channels give rise to a much higher water
permeability of approximately $1.8\times 10^{-2}$ cm/s in
erythrocytes~\cite{YangMaVerkman2001_JBC} while the intrinsic water
permeability of a lipid bilayer membrane is in the range of
$10^{-4}$--$10^{-3}$ cm/s~\cite{ThompsonHuang1966_ANYAS,
FettiplaceHaydon1980_PhysRev, Grafmueller2019_ABLS, Dimova2020_GVB,
BhatiaRobinsonDimova2020_SoftMatt}. 
%
%The swelling/shrinking of a cell is inevitably coupled to changes in
%membrane tension and deformation. Under a strong confinement (such as
%going through a microcapillary or submicron slit),  the transmembrane
%water flow may have comparable contributions from both osmotic and
%mechanical stresses
%\cite{StrokaJiangChenEtAl2014_Cell,LiMoriSun2015_PRL,yao-mor2017}.  
Molecular dynamic (MD) simulations show that a stretched red blood cell
(RBC) membrane porates when the membrane tension reaches the order of
$2$--$4$ mN/m (\cite{RazizadehNikfarPaulLiu2020_BJ} and references
therein), giving rise to enhanced membrane permeability to both water
and macromolecules.  Consequently, as long as the lipid bilayer membrane
remains intact (no poration or rupture), RBCs and vesicles are assumed
to be of the same volume throughout experiments of durations no longer
than tens of minutes.
%
%membrane poration is inevitable when membrane tension is of the order
%of $2-4$ mN/m (\cite{RazizadehNikfarPaulLiu2020_BJ} and references
%therein). 
%%The CGMD results further show that a stretched RBC membrane may have pores that might contribute to transmembrane transport of both water and ATP molecules \cite{RazizadehNikfarPaulLiu2020_BJ}. 
%Thus an important outstanding question in stress-induced ATP release of
%highly deformed RBCs is if there is any transmembrane flow when RBCs
%release ATP as they experience large mechanical load due to confinement
%or large shear stress?  To address such question, the first step we
%take is to focus on the permeable water flux induced by mechanical
%stress.
%%
%%RBCs are known to release ATP under large deformation and microfluidic experiments have been used to study the correlation between RBC deformation and ATP release \cite{Wan2008_PNAS,ForsythWan2011_PNAS,ZhangShenHoganBarakatMisbah2018_BJ}.
%%
%%Why is it interesting to study the dynamics of a semipermeable membrane induced by mechanical stress/confinement?
%%
%%The osmolarity-induced water flux has been well studied, while mechanical pressure may also contribute to the permeable water flow \cite{yao-mor2017}.
%%The response of a cell to an osmotic gradient is inevitably both mechanical (tension redistribution and shape change) and chemical (water flux across the membrane).  
Under what condition is this a good assumption? Do vesicles conserve
volume over hours under flow? What is the dynamic consequence when water
influx/efflux affects the membrane hydrodynamics? To answer these
questions, we investigate the balance between hydrodynamics, tension,
and elasticity of a semipermeable membrane under various flowing
conditions and confinement. 
%Specifically our numerical simulations yield the quantification of the
%time scale and length scale over which water permeation becomes
%important.

% In this work we use numerical simulations to examine the hydrodynamics
% of a semipermeable vesicle in two dimensions. 
In microfluidic experiments, it is challenging to keep a freely
suspended RBC or vesicle in a steady flow for more than a few hours. It
is also difficult to quantify the vesicle area/volume ratio accurately.
Using numerical simulations we can quantify the effects of water
permeability on the hydrodynamics of a vesicle under mechanical
stresses. Vesicles (self-enclosed lipid bilayer membranes) have been
used as a model system to study reshaping, remodeling, and scission of
cell membranes due to osmotic stress~\cite{OgleckaEtAl2014_eLife,
CamposSaric2020_bioRxiv, Dimova2020_GVB, BhatiaChrist2020_SoftMatt}.
Under an osmotic shock, a vesicle may undergo oscillatory variation in
its enclosed volume~\cite{ChabanonHoLiedberg2017_BJ}. When trapped in a
microfluidic channel, an osmotic shock induces vesicle shape changes,
tank-treading and tumbling
motions~\cite{BhatiaRobinsonDimova2020_SoftMatt,
BhatiaChrist2020_SoftMatt}, an illustration that the permeable water
flow is inevitably coupled with the mechanical pressure jump across the
membrane~\cite{yao-mor2017}, which could become comparable to osmotic
stress when the membrane deformation is large~\cite{LiMoriSun2015_PRL,
yao-mor2017}. In the absence of an osmotic gradient, we first show that
a freely suspended semipermeable vesicle behaves differently from an
impermeable vesicle over a long time. Next we show that, at normal
physiological conditions, the permeable water flow can be amplified by
extreme confinement and results in observable change in both vesicle
hydrodynamics and reduced area. These results will shed light on the
dynamic aging process of living cells that constantly go through flows
and constrictions over a period of days.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]
  \centering
  \includegraphics[width=0.5\textwidth]{figures/schematic.pdf}
  \caption{\label{fig:sketch} Schematics showing the effects of
  semipermeability on vesicle hydrodynamics in various configurations.
  The slipper in (a) is the steady-state shape of a semipermeable
  vesicle in a Poiseuille flow.}
\end{figure}

\section*{Formulation}
We consider a two-dimensional semipermeable vesicle (permeable only to
water) suspended in a viscous fluid (Figure~\ref{fig:sketch}). To focus
on the effect of water flow on vesicle hydrodynamics in various
conditions, we assume the fluid is the same inside and outside the
vesicle membrane. On the membrane the permeable water flux gives rise to
a difference between the fluid velocity ${\uu}({\xx}\in \gamma)$ and the
membrane velocity $\dot{\xx}$~\cite{yao-mor2017}:
\begin{align}
  \uu - \dot{\xx} = - k_w (R T \triangle c + \ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $\gamma$ denotes the membrane boundary with an outward normal
$\nn$ (Figure~\ref{fig:sketch}), $k_w$ is the water permeability, $R$ is
the ideal gas constant, $T$ is the temperature ($^{\circ}$K), $\triangle
c$ is contrast in a solute concentration (mol/volume), and
$\ff_\mathrm{mem}$ is the membrane stress that consists of both bending
$ \ff_\mathrm{ben}= -k_b \xx_{ssss}$ and tension
$\ff_\mathrm{ten}=(\sigma \xx_s)_s$ ($s$ is the arclength along the
membrane, see Supplementary Material). Using $k_w\sim 10^{-12}$
m$^2$s/kg (typical of mammalian cell membranes~\cite{LiMoriSun2015_PRL})
and an osmolarity $\triangle c = 0.1$ mol/L (consistent with the osmotic
forces in vesicle aspiration
experiments~\cite{OlbrichRawiczNeedhamEtAl2000_BJ}), the corresponding
water flow is of the order $v_\mathrm{o}=0.1 \mu$m/s. Assuming that the
membrane bending stiffness $k_b = 10^{-19}$J, a membrane tension
$10^{-3}$N/m (one tenth of the lysis tension) and a characteristic
length scale of $10^{-6}$m, the tension contribution ($v_\mathrm{t}$)
and elastic stress contribution ($v_\mathrm{e}$) to the permeable water
flow can be scaled to $v_\mathrm{o}$ as
$v_\mathrm{o}:v_\mathrm{t}:v_\mathrm{e}=10^{0}:10^{-2}:10^{-6}$.
Furthermore, we note that in the absence of an osmotic stress, the water
permeable flow can be significant when the membrane tension is large due
to the external stress such as the hydrodynamics stress due to an
external flow, confinement, or both. In the following we will explore
both.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Effects of membrane water permeability}
We consider a single semipermeable vesicle in familiar configurations to
highlight the effects of membrane water permeability. In free space we
place a semipermeable vesicle in a quiescent flow, a planar shear flow,
and a Poiseuille flow. We also consider a semipermeable vesicle in
confinement: a long channel and a contracting channel. For each example,
we fix the vesicle length and examine the enclosed area as a dynamical
consequence of water flow, vesicle shape deformation, and membrane
tension distribution.

In experiments, vesicles under shear flows are measured for a minute
(chapter 19 in~\cite{Dimova2020_GVB}), a duration over which vesicle can
be assumed impermeable~\cite{AbkarianViallat2005_BJ}. Our simulations of
a semipermeable vesicle in these flows show that water permeability may
lead to different vesicle hydrodynamics over a time scale of 10 to 100
minutes. 
%{\bf YNY: is this correct? The shear flow results seem to suggest that
%the permeability may lead to change in area in less than a second???}
For vesicles in a narrow microfluidic channel, the typical time scale is
milliseconds~\cite{abk-fai-sto2006}. Our simulations show that strong
confinement can lead to amplification of water permeability at such
short time scale.
%
%all vesicles have a length
%$L$, but the area is not conserved because of semipermeability. We vary
%the initial vesicle reduced area, the permeability rate, and the flow
%rate. For all simulations, the bending modulus is $10^{-19}J$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/relaxationComposite.tikz}
  \else
  \includegraphics{figures/relaxationComposite.pdf}
  \fi
  \caption{\label{fig:relaxationComposite} (a): Snapshots of four
  semipermeable vesicles with $\beta = 10^{-3}$ submerged in a quiescent
  flow. The red regions correspond to influx and the blue regions
  correspond to efflux. In all cases, the vesicle reaches a steady state
  circular shape. (b): The reduced area of each of the simulations. The
  particular snapshots in (a) occur at the marks along the curve.}
\end{figure*}
In a quiescent flow, a semipermeable vesicle in free space can be
quantified in terms of its area dynamics and its transition to the
equilibrium shape. Our analysis in Supplementary Material shows that a
semipermeable vesicle relaxes to a circle at a rate that is proportional
to $\beta$, and the steady state tension is $\Lambda = 2\pi^2/L^2$. Our
analysis further shows that the time for a semipermeable vesicle to
deviate from its initial reduced area and to reach its steady state is
proportional to $\beta^{-1}$. For this reason, our numerical results
below are reported in terms of the scaled time $\beta t$.

@@

Between $t=10$ seconds and $t=10$ minutes, the intermediate shapes of
semipermeable vesicles resemble the shape of an impermeable vesicle.
This can be explained by considering the two terms that contribute to
the vesicle velocity: (i) the permeability velocity, $\beta(\ff \cdot
\nn) \nn$, and (ii) the force balance velocity, $\SS[\ff]$. Since $\beta
= 10^{-6}$, the permeability velocity is much smaller than the force
balance velocity at early times, and the semipermeable vesicle dynamics
resembles those of an impermeable vesicle. As the force balance velocity
decreases, the permeability velocity dominates, and the vesicle begins
to inflate to a circle with no vanishing force and velocity. 
\todo[inline]{There might be some confusion regarding the value of
$\beta$. The caption says one value and the text says another.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Planar Shear Flow}
\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/shearComposite.tikz}
  \else
  \includegraphics{figures/shearComposite.pdf}
  \fi
  \caption{\label{fig:shearComposite} (a): The reduced area of a
  semipermeable vesicle with $\beta = 10^{-3}$ initialized with four
  different reduced areas in a shear flow with two different flow rates.
  (b): The steady state shape of a semipermeable vesicle with varying
  flow rates and semipermeability rates. The circled vesicles correspond
  to the simulations in part (a).}
\end{figure}
